---
title: "Notes on Mediterranean Swordfish"
subtitle: "General notes on the swordfish assessment"
author: "Danai"
date: "2026-05-31"
format: html
execute:
  echo: false
  warning: false
  message: false
---
## Input for assessment

```{r}
library(FLCore)
library(FLa4a)
library(FLBRP)
library(ggplot2); theme_set(theme_minimal())
```

### Growth parameters:
VB params: Linf=238.58, k=0.185, t0=-1.404 

### Length weight
a=0.0000096102, b=3.059

### Maturity Vector
```{r}

#| label: tbl-maturity
#| tbl-cap: "Maturity vector for ages 0 to 9."

mat_0_9  <- c(0.00, 0.00, 0.15, 0.65, 1.00, 1.00, rep(1.00, 4))

maturity_table <- rbind(
  Age = paste0("**", 0:9, "**"),      
  Maturity = sprintf("%.2f", mat_0_9)  
)

knitr::kable(
  maturity_table, 
  col.names = NULL,                     
  align = "lcccccccccc"                # Left-aligns row names, centers data
)
```

----

### Catches (since 1950)
```{r}
ca <- read.csv("data/SWOMed_JABBA_catch_10April2026.csv")
ca.flq <- FLQuant(as.vector(ca[,-1]), dimnames = list(age = 'all',year = 1950:2024))
plot(ca.flq)
```

## Rejection of past a4a model

### Key Model Differences and Advice

* **Input Data Timeline:** The surplus production model JABBA utilized the entire catch time series from 1950 to 2018, whereas the age-structured models (XSA and a4a) did not include data prior to 1985.
* **Initial 1950 Stock State:** For the JABBA model, the Group accepted the assumption that relative stock biomass was close to unfished levels (~99% of carrying capacity) in 1950, noting it aligned with expert scientific knowledge of the fishery.
* **1985 Stock Status Discrepancies:** The models showed conflicting results for 1985 (the first year considered in XSA):
    * *XSA (including discards):* Estimated a heavily depleted stock, with SSB at 54.1% of $SSB_{MSY}$ and 9.67% of unfished $SSB_{0}$.
    * *JABBA ('Reference' and 'ASEM'):* Estimated a sustainable stock, with $B/B_{MSY} = 1.59$ and $B/B_{0} = 0.70$.

### Biological Implausibility of Age-Structured Models

The Group concluded that the depletion levels estimated by XSA and a4a for 1985 were biologically implausible due to two major conflicts with catch history:

1. **Catches Below MSY:** Total catches never exceeded 14,000 t prior to 1985 and remained consistently below the MSY thresholds of both models (XSA MSY = 16,300 t; JABBA joint-MSY = 13,325 t). Theoretically, overfishing an underexploited stock requires catches to exceed MSY.
2. **Unrealistic Stock Resilience:** XSA paradoxically predicted that despite the stock being severely depleted in 1985, it could still support catches above MSY from 1986–1989 and sustain catches above 80% of MSY through the 1990s and 2000s.

## Final Decision
Because the a4a model results mirrored the flaws of the XSA model, the Group decided to provide its scientific advice for the Mediterranean swordfish stock based exclusively on the two **JABBA** models.


::: {.callout-important}
## Note
The ratios of $SSB/SSB_{MSY}$ from XSA and a4a are not comparable to the ones coming from JABBA model. The exploitable biomass include ages 1,2 and 3 which are not yet fully mature (see below selectivity and maturity curves). 
:::

```{r, echo = FALSE, message = FALSE, warning = FALSE}
load("2020_files/runs/a4a/MCMC/input4MCMC.RData")

# Discards run
fmod <- ~s(year, k = 17) + s(age, k = 3)
srmod <- ~s(year, k = 15)
qmod <- list(~1,~1, ~s(year, k = 3), ~s(year, k = 3),~1,~s(year, k = 9))
fit <- sca(stk_disc, idx_bio, fmodel = fmod, qmodel = qmod, srmodel = srmod)
a4a.stk <- stk + fit
#plot(a4a.stk)
# SRR

landings.n(a4a.stk)<-catch.n(a4a.stk)
landings.wt(a4a.stk)<-catch.wt(a4a.stk)
landings(a4a.stk)<-catch(a4a.stk)
discards(a4a.stk)<-0
discards.n(a4a.stk)<-0
discards.wt(a4a.stk)<-0

sr0 <- as.FLSR(a4a.stk, model = bevholt)
srr <- fmle(sr0,control=list(trace=0),fixed=list(b=100))
#srr <- fmle(sr0,control=list(trace=0))
#plot(srr)
brp0 <- brp(FLBRP(a4a.stk, sr = srr))
#refpts(brp0)
#ssb(a4a.stk)[,'1985']/refpts(brp0)['msy','ssb']
#ssb(a4a.stk)[,'2018']/refpts(brp0)['msy','ssb']
#stock(a4a.stk)[,'1985']/refpts(brp0)['msy','biomass']

a4a.stks <- stk_disc + simulate(fit, nsim = 1000)


brefpt <- FLPar(BMSY = an(refpts(brp0)['msy','biomass']))
yrefpt <- FLPar(MSY = an(refpts(brp0)['msy','yield']))
frefpt <- FLPar(FMSY = an(refpts(brp0)['msy','harvest']))

p1 <- plot(a4a.stks, metrics = list(Biomass = stock)) +
    geom_flpar(data = brefpt, x = 1990) 
p2 <- plot(a4a.stks, metrics = list(Catch = catch)) +
    geom_flpar(data = yrefpt, x = 1995) 
p3 <- plot(a4a.stks, metrics = list(F = fbar)) +
    geom_flpar(data = frefpt, x = 2000) + ggtitle('F')
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#| layout-ncol: 2
#| fig-cap: 
#|   - "Total Stock Biomass from a4a model"
#|   - "Estimated catches from a4a model"
p1
p2
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
flqs <- FLQuants(sel=catch.sel(brp0)/max(catch.sel(brp0)),mat=mat(brp0))
df <- as.data.frame(flqs)
ggplot(data = df) + geom_line(aes(x = age, y = data, group = qname, color = qname))

#xyplot(data~age,data=catch.sel(brp0), type=c('l', 'p'), ylab="f", ylim=c(0,1.5))
```

