---
title: "a4a stochastic slice biomass index"
author: John Doe
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo  = FALSE)
rm(list = ls())
library(reshape2)
library(FLXSA)
library(ggplot2)
library(ggplotFL)
library(FLCore)
library(mixdist)
library(plyr)
library(data.table)
library(dplyr)
library(FLa4a)
library(FLEDA)
library(FLAssess)
library(FLash)
library(FLBRP) 
library(kobe)
library(ggthemes);theme_set(theme_bw())
setwd("~/pCloudDrive/ICCAT/runs/a4a")
load("~/pCloudDrive/ICCAT/data/swom_mixed_data_bioIndexes_corrected_2008.rda")
iFig=0
```


# Data

## Trimmed time series (from 1985)

[...]


```{r prepare data, include=FALSE}
### Stock object
stk <- swom

### set the units
units(landings(stk)) <- "t"
units(catch(stk)) <- "t"
units(catch.n(stk)) <- ""
units(catch.wt(stk)) <- "t"
units(stock.n(stk)) <- ""
units(stock.wt(stk)) <- "t"
units(mat(stk)) <- ""
units(m(stk)) <- "m"
units(harvest(stk)) <- "f"

catch.n(stk)[catch.n(stk) == 0] <- NA

range(stk, c("minfbar", "maxfbar")) = c(2, 4)

### Biomass index
idx <- u
summary(idx)

idx_bio <- FLIndices()

for (i in 1:length(idx)) {
  dms <- list(age="all", year=range(idx[[i]])["minyear"]:range(idx[[i]])["maxyear"])
  idx_bio[[i]] <- FLIndexBiomass(FLQuant(NA, dimnames=dms))
  index(idx_bio[[i]]) <- quantSums(index(idx[[i]]))
}

names(idx_bio) <- names(idx)


for(i in 1:length(idx_bio)) {
  range(idx_bio[[i]],c("startf","endf")) = c(0,1)
  range(idx_bio[[i]],c("min","max")) = c(2,4)
  units(index(idx_bio[[i]])) <- "t"
}
```

```{r trim, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
stk <- trim(stk, year = 1985:2018)
```


```{r catch.n plot, echo=FALSE, fig.height=7, message=FALSE, warning=FALSE, comment=NA}
### catch @ age
plot(catch.n(stk))
```
**Figure `r iFig=iFig+1;iFig`**. Time series of catch-at-age.\

```{r index plot, echo=FALSE, fig.height=7, message=FALSE, warning=FALSE, comment=NA}
## index plots
plot_idx <- as.data.frame(idx_bio)

ggplot(aes(year, data), data = plot_idx[plot_idx$slot == 'index',c(1,3,8,9)]) +
  geom_line(stat = 'identity')+ facet_grid(cname ~.)
```
**Figure `r iFig=iFig+1;iFig`**. Biomass indices.\

```{r cohort consistency, echo=FALSE, fig.height=7, message=FALSE, warning=FALSE, comment=NA}
temp <- FLIndices()
dms <- list(age=range(stk)["min"]:range(stk)["max"], 
            year=range(stk)["minyear"]:range(stk)["maxyear"])
temp[[1]] <- FLIndex(FLQuant(NA, dimnames=dms))
index(temp[[1]]) <- catch.n(stk)
plot(temp[[1]], type="internal")

```

\newpage

# Results fbar 2 - 4

Simple separable factor model on F results in a possible not positive definite covariance matrix which is an indicator of overfitting. The same stands for the built in tensor model. Catchability model is held constant through the years.
Absence of SR model gave an unresonable low recruitment (90.000). Some runs were tested by removing the spanish LLS, the results were almost identical.

## Two "best" runs with a plus group of 6

```{r stks object, echo=FALSE, warning=FALSE, comment=FALSE}
stks <- FLStocks()
nfit = 0
```

```{r fit1, echo=FALSE, warning=FALSE, comment=FALSE}
fmod <- ~s(year, k = 17) + s(age, k = 3)
srmod <- ~s(year, k = 15)
fit1 <- sca(stk, idx_bio, fmodel = fmod, srmodel = srmod)
a4a.stk1 <- stk + fit1
nfit = nfit +1
stks[[as.character(nfit)]] <- a4a.stk1
```

```{r plot fit 1, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(a4a.stk1)
```
**Figure `r iFig=iFig+1;iFig`**. Results for fit `r nfit`. F model : `r paste0(fmod)`, SRmodel: `r paste0(srmod)`.\

```{r fit2, echo=FALSE, warning=FALSE, comment=FALSE}
fmod <- ~s(year, k = 17) + s(replace(age, age>4, 4), k = 3)
srmod <- ~s(year, k = 15)
fit2 <- sca(stk, idx_bio, fmodel = fmod, srmodel = srmod)
a4a.stk2 <- stk + fit2
nfit = nfit +1
stks[[as.character(nfit)]] <- a4a.stk2
```

```{r plot fit 2, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(a4a.stk2)
```
**Figure `r iFig=iFig+1;iFig`**. Results for fit `r nfit`. F model : `r paste0(fmod)`, SRmodel: `r paste0(srmod)`.\

and the corresponding fishing mortalities:

```{r fmort for fit1 fit2, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
c(wireframe(data ~ age + year, data = as.data.frame(harvest(a4a.stk1)), drape = TRUE,
          screen = list(x = -90, y=-45),
          col.regions = colorRampPalette(c("blue", "pink",'yellow'))(100)),
  wireframe(data ~ age + year, data = as.data.frame(harvest(a4a.stk2)), drape = TRUE,
          screen = list(x = -90, y=-45),
          col.regions = colorRampPalette(c("blue", "pink",'yellow'))(100)))
```
**Figure `r iFig=iFig+1;iFig`**. Harvest for fit1 (left) and fit2(right).\

## Diagnostics (bad fit in catch numbers in age 6)

```{r res fit 1, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
res <- residuals(fit1, stk, idx_bio)
plot(res)
```
**Figure `r iFig=iFig+1;iFig`**. Residuals for fit 1.\

```{r res fit 2, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
res <- residuals(fit2, stk, idx_bio)
plot(res)
```
**Figure `r iFig=iFig+1;iFig`**. Residuals for fit 2.\

## Increase by one the knots on splines for age in fishing mortality model


```{r fit3, echo=FALSE, warning=FALSE, comment=FALSE}
fmod <- ~s(year, k = 17) + s(age, k = 4)
srmod <- ~s(year, k = 15)
fit3 <- sca(stk, idx_bio, fmodel = fmod, srmodel = srmod)
a4a.stk3 <- stk + fit3
nfit = nfit +1
stks[[as.character(nfit)]] <- a4a.stk3
```

```{r plot fit 3, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(a4a.stk3)
```
**Figure `r iFig=iFig+1;iFig`**. Results for fit `r nfit`. F model : `r paste0(fmod)`, SRmodel: `r paste0(srmod)`.\

```{r fmort fit3,  echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
wireframe(data ~ age + year, data = as.data.frame(harvest(a4a.stk3)), drape = TRUE, screen = list(x = -90, y=-45),
          col.regions = colorRampPalette(c("blue", "pink",'yellow'))(100))
```
**Figure `r iFig=iFig+1;iFig`**. Fishing mortality for fit `r nfit`

```{r res fit 3, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
res <- residuals(fit3, stk, idx_bio)
plot(res)
```
**Figure `r iFig=iFig+1;iFig`**. Residuals for fit 3.\

```{r plot fit 123, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(stks)
```
**Figure `r iFig=iFig+1;iFig`**. Results for the 3 fits.\

## Set plus group 5.
```{r}
stk2 <- setPlusGroup(stk, 5)
```


```{r fit4, echo=FALSE, warning=FALSE, comment=FALSE}
fmod <- ~s(year, k = 17) + s(age, k = 3)
srmod <- ~s(year, k = 15)
fit4 <- sca(stk2, idx_bio, fmodel = fmod, srmodel = srmod)
a4a.stk4 <- stk2 + fit4
nfit = nfit +1
stks[[as.character(nfit)]] <- a4a.stk4
```

```{r plot fit 4, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(a4a.stk4)
```
**Figure `r iFig=iFig+1;iFig`**. Results for fit `r nfit`. F model : `r paste0(fmod)`, SRmodel: `r paste0(srmod)`.\

```{r fmort fit4,  echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
wireframe(data ~ age + year, data = as.data.frame(harvest(a4a.stk4)), drape = TRUE, screen = list(x = -90, y=-45),
          col.regions = colorRampPalette(c("blue", "pink",'yellow'))(100))
```
**Figure `r iFig=iFig+1;iFig`**. Fishing mortality for fit `r nfit`

```{r res fit 4, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
res <- residuals(fit4, stk2, idx_bio)
plot(res)
```
**Figure `r iFig=iFig+1;iFig`**. Residuals for `r nfit`.\

## same f after age 4

```{r fit5, echo=FALSE, warning=FALSE, comment=FALSE}
fmod <- ~s(year, k = 17) + s(replace(age, age>4, 4), k = 3)
srmod <- ~s(year, k = 15)
fit5 <- sca(stk2, idx_bio, fmodel = fmod, srmodel = srmod)
a4a.stk5 <- stk2 + fit5
nfit = nfit +1
stks[[as.character(nfit)]] <- a4a.stk5
```

```{r plot fit 5, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(a4a.stk5)
```
**Figure `r iFig=iFig+1;iFig`**. Results for fit `r nfit`. F model : `r paste0(fmod)`, SRmodel: `r paste0(srmod)`.\

```{r fmort fit5,  echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
wireframe(data ~ age + year, data = as.data.frame(harvest(a4a.stk5)), drape = TRUE, screen = list(x = -90, y=-45),
          col.regions = colorRampPalette(c("blue", "pink",'yellow'))(100))
```
**Figure `r iFig=iFig+1;iFig`**. Fishing mortality for fit `r nfit`

```{r res fit 5, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
res <- residuals(fit5, stk2, idx_bio)
plot(res)
```
**Figure `r iFig=iFig+1;iFig`**. Residuals for `r nfit`.\


```{r plot fit 45, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
stks_gr5 <- FLStocks(catch_Obs = stk2, fit4 = stks[[4]], fit5 = stks[[5]])
plot(stks_gr5)
```
**Figure `r iFig=iFig+1;iFig`**. Results for the fits with plus group 5.\