---
title: "a4a stochastic slice biomass index"
author: John Doe
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo  = FALSE)
rm(list = ls())
library(reshape2)
library(FLXSA)
library(ggplot2)
library(ggplotFL)
library(FLCore)
library(mixdist)
library(plyr)
library(data.table)
library(dplyr)
library(FLa4a)
library(FLEDA)
library(FLAssess)
library(FLash)
library(FLBRP) 
library(kobe)
library(ggthemes);theme_set(theme_bw())
setwd("~/pCloudDrive/ICCAT/runs/a4a")
load("~/pCloudDrive/ICCAT/data/swom_mixed_data_bioIndexes_corrected_2008.rda")
iFig=0
```

# Intro

[...]

# Materials and Methods

[...]

## Data

```{r prepare data, include=FALSE}
### Stock object
stk <- swom

### set the units
units(landings(stk)) <- "t"
units(catch(stk)) <- "t"
units(catch.n(stk)) <- ""
units(catch.wt(stk)) <- "t"
units(stock.n(stk)) <- ""
units(stock.wt(stk)) <- "t"
units(mat(stk)) <- ""
units(m(stk)) <- "m"
units(harvest(stk)) <- "f"

catch.n(stk)[catch.n(stk) == 0] <- 0.1

range(stk, c("minfbar", "maxfbar")) = c(2, 4)

### Biomass index
idx <- u
summary(idx)

idx_bio <- FLIndices()

for (i in 1:length(idx)) {
  dms <- list(age="all", year=range(idx[[i]])["minyear"]:range(idx[[i]])["maxyear"])
  idx_bio[[i]] <- FLIndexBiomass(FLQuant(NA, dimnames=dms))
  index(idx_bio[[i]]) <- quantSums(index(idx[[i]]))
}

names(idx_bio) <- names(idx)


for(i in 1:length(idx_bio)) {
  range(idx_bio[[i]],c("startf","endf")) = c(0,1)
  range(idx_bio[[i]],c("min","max")) = c(1,3)
  units(index(idx_bio[[i]])) <- "t"
}
```

```{r catch.n plot, echo=FALSE, fig.height=7, message=FALSE, warning=FALSE, comment=NA}
### catch @ age
plot(catch.n(stk))
```
**Figure `r iFig=iFig+1;iFig`**. Time series of catch-at-age.\

```{r index plot, echo=FALSE, fig.height=7, message=FALSE, warning=FALSE, comment=NA}
## index plots
plot_idx <- as.data.frame(idx_bio)

ggplot(aes(year, data), data = plot_idx[plot_idx$slot == 'index',c(1,3,8,9)]) +
  geom_line(stat = 'identity')+ facet_grid(cname ~.)
```
**Figure `r iFig=iFig+1;iFig`**. Biomass indices.\

# Results

```{r stks object, echo=FALSE, warning=FALSE, comment=FALSE}
stks <- FLStocks()
nfit = 0
```


## Full time series

The simple separable model, treating years as categorical variables gave unresonable results in both recruitment and f rates.

Without restraining the fishing mortality model to be the same after age 4 we get a very low f and a very high ssb, variants of the separable splines modeling  on year gives pretty much the same results. (varying k around the half of the range of time series). On the other hand the varying degrees of freedom in smoothing splines on age, gives a completely differrent perception of the stock, which suggests a quite unstable model. 

```{r fit1, echo=FALSE, warning=FALSE, comment=FALSE}
fmod <- ~s(year, k = 23) + s(age, k = 4)
fit1 <- sca(stk, idx_bio, fmodel = fmod)
a4a.stk1 <- stk + fit1
nfit = nfit +1
stks[[as.character(nfit)]] <- a4a.stk1
```

```{r plot fit 1, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(a4a.stk1)
```
**Figure `r iFig=iFig+1;iFig`**. Results for f: `r paste0(fmod)`.\

```{r fit2, echo=FALSE, warning=FALSE, comment=FALSE}
fmod <- ~s(year, k = 23) + s(age, k = 3)
fit2 <- sca(stk, idx_bio, fmodel = fmod)
a4a.stk2 <- stk + fit2
nfit = nfit +1
stks[[as.character(nfit)]] <- a4a.stk2
```

```{r plot fit 2, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(a4a.stk2)
```
**Figure `r iFig=iFig+1;iFig`**. Results for f: `r paste0(fmod)`.\

and the corresponding fishing mortalities:

```{r fmort for fit1 fit2, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
c(wireframe(data ~ age + year, data = as.data.frame(harvest(a4a.stk1)), drape = TRUE,
          screen = list(x = -90, y=-45),
          col.regions = colorRampPalette(c("blue", "pink",'yellow'))(100)),
  wireframe(data ~ age + year, data = as.data.frame(harvest(a4a.stk2)), drape = TRUE,
          screen = list(x = -90, y=-45),
          col.regions = colorRampPalette(c("blue", "pink",'yellow'))(100)))
```
**Figure `r iFig=iFig+1;iFig`**. Harvest for fit1 (left) and fit2(right).\

## Diagnostics

```{r res fit 1, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
res <- residuals(fit1, stk, idx_bio)
plot(res)
```
**Figure `r iFig=iFig+1;iFig`**. Residuals for fit 1.\

```{r res fit 2, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
res <- residuals(fit2, stk, idx_bio)
plot(res)
```
**Figure `r iFig=iFig+1;iFig`**. Residuals for fit 2.\

## Forcing f to be the same after age 4


```{r fit3, echo=FALSE, warning=FALSE, comment=FALSE}
fmod <-  ~ s(replace(age, age>4, 4), k = 4) + s(year, k=23)
fit3 <- sca(stk, idx_bio, fmodel = fmod)
a4a.stk3 <- stk + fit3
nfit = nfit +1
stks[[as.character(nfit)]] <- a4a.stk3
```

```{r plot fit 3, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(a4a.stk3)
```
**Figure `r iFig=iFig+1;iFig`**. Results for f: `r paste0(fmod)`.\

```{r fmort for fit3, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
wireframe(data ~ age + year, data = as.data.frame(harvest(a4a.stk3)), drape = TRUE,
          screen = list(x = -90, y=-45),
          col.regions = colorRampPalette(c("blue", "pink",'yellow'))(100))

```
**Figure `r iFig=iFig+1;iFig`**. Harvest for fit3

```{r res fit 3, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
res <- residuals(fit3, stk, idx_bio)
plot(res)
```
**Figure `r iFig=iFig+1;iFig`**. Residuals for fit 3.\

\newpage

```{r plot fit 123, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(stks)
```
**Figure `r iFig=iFig+1;iFig`**. Results for the 3 fits.\


Add a smoother to the last run for SR $s(year, k =20)$:

```{r fit4, echo=FALSE, warning=FALSE, comment=FALSE}
fmod <-  ~ s(replace(age, age>4, 4), k = 4) + s(year, k=23)
srmod <- ~s(year, k = 20)
fit4 <- sca(stk, idx_bio, fmodel = fmod, srmodel = srmod)
a4a.stk4 <- stk + fit4
nfit = nfit +1
stks[[as.character(nfit)]] <- a4a.stk4
```

```{r plot fit 4, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(a4a.stk4)
```
**Figure `r iFig=iFig+1;iFig`**. Results for f: `r paste0(fmod)`.\


```{r res fit 4, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
res <- residuals(fit4, stk, idx_bio)
plot(res)
```
**Figure `r iFig=iFig+1;iFig`**. Residuals for fit 4.\

```{r plot fit 3 n 4, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(FLStocks(catch_obs = stk, fit3 = stks[['3']], fit4 = stks[['4']]))
```
**Figure `r iFig=iFig+1;iFig`**. Results for fit3 and fit4.\

```{r retro for fit 4, echo=FALSE, warning=FALSE, comment=FALSE}
retro <- function(stk, idxs, retro=3, kfrac="missing", k, ...){
  args <- list(...)
  if(missing(kfrac)) kfrac <- 0.5
  lst0 <- split(0:retro, 0:retro)
  lst0 <- lapply(lst0, function(x){
    yr <- range(stk)["maxyear"] - x
    args$stock <- window(stk, end=yr)
    args$indices <- FLIndices(window(idxs, end=yr))
    KY <- unname(k - floor(x*kfrac))
    # NOTE THIS HAS TO BE ADAPTED FOR THE MODEL USED  !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
    fmod <- substitute(~ s(replace(age, age>4, 4), k = 4) + s(year, k=KY), list(KY=KY))
    args$fmodel <- as.formula(fmod)
    fit <- do.call("sca", args)
    args$stock + fit
  })
  FLStocks(lst0)
}
```

```{r plot retro for fit4, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
retro.lst <- retro(stk, idx_bio, k=23, srmodel= srmod)
plot(retro.lst, main="Retrospective analysis")
```
**Figure `r iFig=iFig+1;iFig`**. Retrospective plot for fit4.\



