---
title: "a4a stochastic slice biomass index"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo  = FALSE)
rm(list = ls())
library(reshape2)
library(FLXSA)
library(ggplot2)
library(ggplotFL)
library(FLCore)
library(mixdist)
library(plyr)
library(data.table)
library(dplyr)
library(FLa4a)
library(FLEDA)
library(FLAssess)
library(FLash)
library(FLBRP) 
library(kobe)
library(ggthemes);theme_set(theme_bw())
setwd("~/pCloudDrive/ICCAT/runs/a4a")
load("~/pCloudDrive/ICCAT/data/swom_mixed_data_bioIndexes_corrected_2008.rda")
iFig=0
```


# Data

## Trimmed time series (from 1985)

[...]


```{r prepare data, include=FALSE}
### Stock object
stk <- swom

### set the units
units(landings(stk)) <- "t"
units(catch(stk)) <- "t"
units(catch.n(stk)) <- ""
units(catch.wt(stk)) <- "t"
units(stock.n(stk)) <- ""
units(stock.wt(stk)) <- "t"
units(mat(stk)) <- ""
units(m(stk)) <- "m"
units(harvest(stk)) <- "f"

catch.n(stk)[catch.n(stk) == 0] <- 0.1

range(stk, c("minfbar", "maxfbar")) = c(2, 4)

### Biomass index
idx <- u
summary(idx)

idx_bio <- FLIndices()

for (i in 1:length(idx)) {
  dms <- list(age="all", year=range(idx[[i]])["minyear"]:range(idx[[i]])["maxyear"])
  idx_bio[[i]] <- FLIndexBiomass(FLQuant(NA, dimnames=dms))
  index(idx_bio[[i]]) <- quantSums(index(idx[[i]]))
}

names(idx_bio) <- names(idx)


for(i in 1:length(idx_bio)) {
  range(idx_bio[[i]],c("startf","endf")) = c(0,1)
  range(idx_bio[[i]],c("min","max")) = c(2,4)
  units(index(idx_bio[[i]])) <- ""
}
```

```{r trim, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
stk <- trim(stk, year = 1985:2018)
```

```{r Lorenzen M, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
stk_L <- stk
par=FLPar(Linf=238.59,K=0.185,t0=-1.404,a=9.62e-3,b=3.06,l50=132,
          m1=20/c(mean(stock.wt(stk_L)["3"])),m2=FLPar(-0.309))

m=FLife:::lorenzen(stock.wt(stk_L),par[c("m1","m2")])
par["m1"]=0.2*par["m1"]/mean(m["3"])  

m(stk_L)=FLife:::lorenzen(stock.wt(stk_L),par[c("m1","m2")])
```

```{r catch.n plot, echo=FALSE, fig.height=7, message=FALSE, warning=FALSE, comment=NA}
### catch @ age
plot(catch.n(stk))
```
**Figure `r iFig=iFig+1;iFig`**. Time series of catch-at-age.\

```{r index plot, echo=FALSE, fig.height=7, message=FALSE, warning=FALSE, comment=NA}
## index plots
plot_idx <- as.data.frame(idx_bio)

ggplot(aes(year, data), data = plot_idx[plot_idx$slot == 'index',c(1,3,8,9)]) +
  geom_line(stat = 'identity')+ facet_grid(cname ~.)
```
**Figure `r iFig=iFig+1;iFig`**. Biomass indices.\

```{r cohort consistency, echo=FALSE, fig.height=7, message=FALSE, warning=FALSE, comment=NA}
temp <- FLIndices()
dms <- list(age=range(stk)["min"]:range(stk)["max"], 
            year=range(stk)["minyear"]:range(stk)["maxyear"])
temp[[1]] <- FLIndex(FLQuant(NA, dimnames=dms))
index(temp[[1]]) <- catch.n(stk)
plot(temp[[1]], type="internal")

```

\newpage

# Results fbar 2 - 4

Simple separable factor model on F results in a possible not positive definite covariance matrix which is an indicator of overfitting. The same stands for the built in tensor model. Catchability model is held constant through the years.
Absence of SR model gave an unresonable low recruitment (90.000). Some runs were tested by removing the spanish LLS, the results were almost identical.

## Natural mortality 0.2

```{r stks object, echo=FALSE, warning=FALSE, comment=FALSE}
stks <- FLStocks()
nfit = 0
```

```{r fit1, echo=FALSE, warning=FALSE, comment=FALSE}
fmod <- ~s(year, k = 17) + s(age, k = 3)
srmod <- ~s(year, k = 17)
fit1 <- sca(stk, idx_bio, fmodel = fmod, srmodel = srmod)
a4a.stk1 <- stk + fit1
nfit = nfit +1
stks[[as.character(nfit)]] <- a4a.stk1
```

```{r plot fit 1, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(a4a.stk1)
```
**Figure `r iFig=iFig+1;iFig`**. Results for fit `r nfit`. F model : `r paste0(fmod)`, SRmodel: `r paste0(srmod)`.\

## Lorenzen Natural Mortality

```{r fit2, echo=FALSE, warning=FALSE, comment=FALSE}
fmod <- ~s(year, k = 17) + s(age, k = 3)
srmod <- ~s(year, k = 15)
fit2 <- sca(stk_L, idx_bio, fmodel = fmod, srmodel = srmod)
a4a.stk2 <- stk_L + fit2
nfit = nfit +1
stks[[as.character(nfit)]] <- a4a.stk2
```

```{r plot fit 2, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(a4a.stk2)
```
**Figure `r iFig=iFig+1;iFig`**. Results for fit `r nfit`. F model : `r paste0(fmod)`, SRmodel: `r paste0(srmod)`.\

and the corresponding fishing mortalities:

```{r fmort for fit1 fit2, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
c(wireframe(data ~ age + year, data = as.data.frame(harvest(a4a.stk1)), drape = TRUE,
          screen = list(x = -90, y=-45), zlab = 'F',
          col.regions = colorRampPalette(c("blue", "pink",'yellow'))(100)),
  wireframe(data ~ age + year, data = as.data.frame(harvest(a4a.stk2)), drape = TRUE,
          screen = list(x = -90, y=-45), zlab = 'F',
          col.regions = colorRampPalette(c("blue", "pink",'yellow'))(100)))
```
**Figure `r iFig=iFig+1;iFig`**. Harvest for M = 0.2 (left) and Lorenzen (right).\

## Diagnostics (bad fit in catch numbers in age 6)

```{r res fit 1, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
res <- residuals(fit1, stk, idx_bio)
plot(res)
```
**Figure `r iFig=iFig+1;iFig`**. Residuals for fit 1.\

```{r res fit 2, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
res <- residuals(fit2, stk_L, idx_bio)
plot(res)
```
**Figure `r iFig=iFig+1;iFig`**. Residuals for fit 2.\


```{r plot fit 123, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE, comment=NA}
plot(stks)
```
**Figure `r iFig=iFig+1;iFig`**. Results for the 2 fits.\

```{r reference points}
##### Reference points

landings.n(a4a.stk2)<-catch.n(a4a.stk2)
landings.wt(a4a.stk2)<-catch.wt(a4a.stk2)
landings(a4a.stk2)<-catch(a4a.stk2)
discards(a4a.stk2)<-0
discards.n(a4a.stk2)<-0
discards.wt(a4a.stk2)<-0

model = "bevholt"
spe_srr <- fmle(as.FLSR(a4a.stk2, model = model))
spe_brp <- brp(FLBRP(a4a.stk2, sr = spe_srr))
ref_points<- refpts(spe_brp)
ref_points
plot(spe_brp, obs = T)


ple4SR <- transform(as.FLSR(a4a.stk2, model=bevholt), ssb=ssb/100, rec=rec/100)
ple4SR <- fmle(ple4SR,control=list(silent=T))
params(ple4SR)['b',] <- params(ple4SR)['b',] / 100
ple4SR <- transform(ple4SR, ssb=ssb*100, rec=rec*100)
brp4Ri <- FLBRP(a4a.stk2, sr=ple4SR)
plot(brp4Ri, obs = T)
# stk.brp <- brp(FLBRP(a4a.stk2))
# plot(stk.brp)
# refpts(stk.brp)['f0.1','harvest']
```

