---
title: Final Assessment Runs using a4a
author: "D. Mantopoulou-Palouka and G. Tserpes"
output:
  ioslides_presentation:
    smaller: yes
    widescreen: yes
  html_document:
    df_print: paged
  beamer_presentation: default
---

<style>
.forceBreak { -webkit-column-break-after: always; break-after: column; }
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.height = "90%")
rm(list = ls())
library(reshape2)
library(readxl)
# library(FLXSA)
library(ggplot2)
library(ggplotFL)
library(FLCore)
library(mixdist)
library(plyr)
library(data.table)
library(dplyr)
library(FLa4a)
library(FLEDA)
library(FLAssess)
library(FLash)
library(FLBRP) 
library(kobe)
library(knitr)
library(scales)
library(ggthemes);theme_set(theme_bw())
setwd("~/pCloudDrive/ICCAT/runs/a4a")
load("~/pCloudDrive/ICCAT/data/swom_mixed_data_bioIndexes_corrected_2008.rda")
iFig=0
iTab=0


### Stock object
stk <- swom

### Index object
idx <- u

### Stock object with discards
load("~/pCloudDrive/ICCAT/data/swom_mixed_data_bioIndexes_dsc.rda")
stk_disc <- swom

### set the units
units(landings(stk)) <- units(landings(stk_disc)) <-"t"
units(catch(stk)) <- units(catch(stk_disc)) <- "t"
units(catch.n(stk)) <- units(catch.n(stk_disc)) <- ""
units(catch.wt(stk)) <- units(catch.wt(stk_disc)) <- "t"
units(stock.n(stk)) <- units(stock.n(stk_disc)) <- ""
units(stock.wt(stk)) <- units(stock.wt(stk_disc)) <- "t"
units(mat(stk)) <- units(mat(stk_disc)) <- ""
units(m(stk)) <- units(m(stk_disc)) <- "m"
units(harvest(stk)) <- units(harvest(stk_disc)) <- "f"

# catch.n(stk)[catch.n(stk_disc) == 0] <- 0.1

range(stk, c("minfbar", "maxfbar")) = c(2, 4)
range(stk_disc, c("minfbar", "maxfbar")) = c(2, 4)

### Biomass index
idx <- u
summary(idx)

idx_bio <- FLIndices()

for (i in 1:length(idx)) {
  dms <- list(age="all", year=range(idx[[i]])["minyear"]:range(idx[[i]])["maxyear"])
  idx_bio[[i]] <- FLIndexBiomass(FLQuant(NA, dimnames=dms))
  index(idx_bio[[i]]) <- quantSums(index(idx[[i]]))
}

names(idx_bio) <- names(idx)


for(i in 1:length(idx_bio)) {
  range(idx_bio[[i]],c("startf","endf")) = c(0,1)
  range(idx_bio[[i]],c("min","max")) = c(2,4)
  units(index(idx_bio[[i]])) <- ""
}


lig_sur <- read_xlsx("~/pCloudDrive/ICCAT/data/indices.xlsx")
lig_sur <- lig_sur[,c(1,13)]
names(lig_sur) <- c('year', 'LIG_SUR')
lig_sur <- lig_sur[complete.cases(lig_sur), ]

dms <- list(age="all", year=min(lig_sur$year):max(lig_sur$year))

index_lig <- FLQuant(lig_sur$LIG_SUR,dimnames = dms)
lig_sur_idx <- FLIndexBiomass(FLQuant(NA, dimnames=dms))
index(lig_sur_idx) <- index_lig
range(lig_sur_idx,c("startf","endf")) = c(0,1)
range(lig_sur_idx,c("min","max")) = c(2,4)
units(index(lig_sur_idx)) <- ""

idx_bio[[6]] <- lig_sur_idx
names(idx_bio) <- c("GR_LL","SP_LL","MO_LL","LI_LL","SI_LL" ,"LI_SUR")


### Trim the stocks
stk <- trim(stk, year = 1985:2018)
stk_disc <- trim(stk_disc, year = 1985:2018)

### Set plus group
stk <- setPlusGroup(stk, 5)
stk_disc <- setPlusGroup(stk_disc, 5)

```
## Data input for the assessment

### Catch Data
- Catch-at-age data since 1985
- plus group 5
- fbar at 2 - 4

### CPUE Data

- Six standardized CPUE series

    - Greek longliners since 1987
    - Moroccan longliners since 2012
    - Spanish longliners since 1988
    - Ligurian longliners since 2010
    - Sicilian longliners from 1991 to 2009
    - Ligurian surface longliners 1991 to 2009
    
- Representative of age groups 2 - 4

## CPUE indices

```{r index plot, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
## index plots
plot_idx <- as.data.frame(idx_bio)

ggplot(aes(year, data), data = plot_idx[plot_idx$slot == 'index',c(1,3,8,9)]) +
  geom_line(stat = 'identity')+ facet_grid(cname ~.)+ggtitle("Time series of standardized CPUE indices")
```


## Final runs

- Two final runs were considered:

    - one with constant M 
    - one with constant M including estimated Discards

- Models specifications kept the same:
    
    - for $F$, $R$, $\mathbf{\sigma^2}$, $\tau^{2}_{s}$ and $N_{1}$

    - Q model altered according to the new CPUE index:
    
$$  
Q : 
     \begin{cases}
       \sim 1, &\quad \text{Greek longline}\\
       \sim 1, &\quad \text{Spanish longline}\\
       \sim s(year, k=3), &\quad \text{Moroccan longline}\\
       \sim s(year, k=3), &\quad \text{Ligurian longline}\\
       \sim 1, &\quad \text{Sicilian longline}\\
       \sim s(year, k=9), &\quad \text{Ligurian surface longline}\\
     \end{cases}
$$


## final run | assuming constant M

```{r RUN constM, comment=NA, message=F, echo=F, warning=F}
fmod2 <- ~s(year, k = 17) + s(age, k = 3)
srmod1 <- ~s(year, k = 15)
srmod2 <- ~bevholt(CV = 0.2)
qmod3 <- list(~1,~1, ~s(year, k = 3), ~s(year, k = 3),~1,~s(year, k = 9))
fit9 <- sca(stk, idx_bio, fmodel = fmod2, srmodel = srmod1, qmodel = qmod3)
a4a.stk9 <- stk + fit9
```

```{r plot constM, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
fits9 <- simulate(fit9, 1000)
a4a.stk9_sim <- stk + fits9
plot(a4a.stk9_sim)
```



## Diagnostics | final run assuming constant M

```{r res}
res <- residuals(fit9, stk, idx_bio)
res_Df <- as.data.frame(res)
res_Df <- res_Df[,c(1,2,7,8)]
```

Standardized residuals of the catch numbers by age
```{r std res constM, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
# standardized residuals
ggplot(aes(year,data), data = res_Df[res_Df$qname == "catch.n",]) + geom_point() + facet_wrap(age ~.) + geom_smooth()
```

## Diagnostics | final run assuming constant M

Standardized residuals of the CPUE series
```{r std res idx constM, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
# standardized residuals
ggplot(aes(year,data), data = res_Df[res_Df$qname != "catch.n"&res_Df$qname != "catch",]) +
  geom_point() + facet_wrap(qname ~.) + geom_smooth()

```

## Retrospective plot | final run assuming constant M

```{r retro constM, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
retro <- function(stk, idxs, retro=2, kfrac="missing", k, k2, ...){
  args <- list(...)
  if(missing(kfrac)) kfrac <- 0.5
  lst0 <- split(0:retro, 0:retro)
  lst0 <- lapply(lst0, function(x){
    yr <- range(stk)["maxyear"] - x
    args$stock <- window(stk, end=yr)
    args$indices <- FLIndices(window(idxs, end=yr))
    KY <- unname(k - floor(x*kfrac))
    KZ <- unname(k2 - floor(x*kfrac))
    # NOTE THIS HAS TO BE ADAPTED FOR THE MODEL USED  !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
    fmod <- substitute(~s(year, k = KY) + s(age, k = 3), list(KY=KY))
    srmod <- substitute(~s(year, k = KZ), list(KZ=KZ))
    # fmod <- substitute(~s(year, k = KY) + s(replace(age, age>4, 4), k = 3), list(KY=KY))
    args$fmodel <- as.formula(fmod)
    args$srmodel <- as.formula(srmod)
    fit <- do.call("sca", args)
    args$stock + fit
  })
  FLStocks(lst0)
}
retro.lst <- retro(stk, idx_bio, k=17, k2 = 15, qmodel = qmod3)
plot(retro.lst)#+theme(legend.position="none")
```

## final run | assuming constant M discards included

```{r Run constM discards, include=FALSE}
############## Discards included and 6 CPUE indices ######################## 
fmod2 <- ~s(year, k = 17) + s(age, k = 3)
srmod1 <- ~s(year, k = 15)
# qmod2 <- list(~1,~1, ~s(year, k = 3), ~s(year, k = 3),~1,~1)
qmod3 <- list(~1,~1, ~s(year, k = 3), ~s(year, k = 3),~1,~s(year, k = 9))
fit9_disc <- sca(stk_disc, idx_bio, fmodel = fmod2, srmodel = srmod1, qmodel = qmod3)
a4a.stk9_disc <- stk_disc + fit9_disc
```

```{r plot constM disc, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
fits9_disc <- simulate(fit9_disc, 1000)
a4a.stk9_disc_sim <- stk_disc + fits9_disc
plot(a4a.stk9_disc_sim)
```

## Diagnostics | final run assuming constant M discards included

```{r res disc}
res <- residuals(fit9_disc, stk_disc, idx_bio)
res_Df <- as.data.frame(res)
res_Df <- res_Df[,c(1,2,7,8)]
```

Standardized residuals of the catch numbers by age
```{r std res constM disc, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
# standardized residuals
ggplot(aes(year,data), data = res_Df[res_Df$qname == "catch.n",]) + geom_point() + facet_wrap(age ~.) + geom_smooth()
```

## Diagnostics | final run assuming constant M discards included

Standardized residuals of the CPUE series
```{r std res idx constM disc, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
# standardized residuals
ggplot(aes(year,data), data = res_Df[res_Df$qname != "catch.n"&res_Df$qname != "catch",]) +
  geom_point() + facet_wrap(qname ~.) + geom_smooth()
```

## Retrospective plot | final run assuming constant M discards included

```{r retro constM disc, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
retro <- function(stk, idxs, retro=2, kfrac="missing", k, k2, ...){
  args <- list(...)
  if(missing(kfrac)) kfrac <- 0.5
  lst0 <- split(0:retro, 0:retro)
  lst0 <- lapply(lst0, function(x){
    yr <- range(stk)["maxyear"] - x
    args$stock <- window(stk, end=yr)
    args$indices <- FLIndices(window(idxs, end=yr))
    KY <- unname(k - floor(x*kfrac))
    KZ <- unname(k2 - floor(x*kfrac))
    # NOTE THIS HAS TO BE ADAPTED FOR THE MODEL USED  !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
    fmod <- substitute(~s(year, k = KY) + s(age, k = 3), list(KY=KY))
    srmod <- substitute(~s(year, k = KZ), list(KZ=KZ))
    # fmod <- substitute(~s(year, k = KY) + s(replace(age, age>4, 4), k = 3), list(KY=KY))
    args$fmodel <- as.formula(fmod)
    args$srmodel <- as.formula(srmod)
    fit <- do.call("sca", args)
    args$stock + fit
  })
  FLStocks(lst0)
}
retro.lst <- retro(stk_disc, idx_bio, k=17, k2 = 15, qmodel = qmod3)
plot(retro.lst)#+theme(legend.position="none")
```


## Comparison plot

```{r reference points constM, message=FALSE, warning=FALSE, comment=NA, include=FALSE}
##### Reference points

landings.n(a4a.stk9)<-catch.n(a4a.stk9)
landings.wt(a4a.stk9)<-catch.wt(a4a.stk9)
landings(a4a.stk9)<-catch(a4a.stk9)
discards(a4a.stk9)<-0
discards.n(a4a.stk9)<-0
discards.wt(a4a.stk9)<-0

model = "bevholt"
srr <- fmle(as.FLSR(a4a.stk9, model = model),control=list(trace=0),fixed=list(b=100))
brp <- brp(FLBRP(a4a.stk9, sr = srr))
ref_points<- refpts(brp)
```

```{r reference points disc, message=FALSE, warning=FALSE, comment=NA, include=FALSE}

landings.n(a4a.stk9_disc)<-catch.n(a4a.stk9_disc)
landings.wt(a4a.stk9_disc)<-catch.wt(a4a.stk9_disc)
landings(a4a.stk9_disc)<-catch(a4a.stk9_disc)
discards(a4a.stk9_disc)<-0
discards.n(a4a.stk9_disc)<-0
discards.wt(a4a.stk9_disc)<-0

model = "bevholt"
srr_disc <- fmle(as.FLSR(a4a.stk9_disc, model = model),control=list(trace=0),fixed=list(b=100))
brp_disc <- brp(FLBRP(a4a.stk9_disc, sr = srr_disc))
ref_points_disc <- refpts(brp_disc)
```

```{r ref total df, include=FALSE}
tot_ref <- data.frame(value = c('Fmsy','Bmsy','F2018','SSB','FFmsy','FF0.1','BBmsy'),
                      constantM = c(ref_points['msy','harvest'],
                                    ref_points['msy','biomass'],
                                    fbar(a4a.stk9)[[34]],
                                    ssb(a4a.stk9)[[34]],
                                    fbar(a4a.stk9)[[34]]/ref_points['msy','harvest'],
                                    fbar(a4a.stk9)[[34]]/ref_points['f0.1','harvest'],
                                    stock(a4a.stk9)[[34]]/ref_points['msy','biomass']),
                      Disc_constM = c(ref_points_disc['msy','harvest'],
                                    ref_points_disc['msy','biomass'],
                                    fbar(a4a.stk9_disc)[[34]],
                                    ssb(a4a.stk9_disc)[[34]],
                                    fbar(a4a.stk9_disc)[[34]]/ref_points_disc['msy','harvest'],
                                    fbar(a4a.stk9_disc)[[34]]/ref_points_disc['f0.1','harvest'],
                                    stock(a4a.stk9_disc)[[34]]/ref_points_disc['msy','biomass'])
                      )


```

```{r plot stocks, echo=FALSE}
plot(FLStocks(Constant_M = a4a.stk9_sim, 
              Discards = a4a.stk9_disc_sim))
```

<p class="forceBreak"></p>

## Reference points

```{r ref points table, echo=FALSE}
kable(tot_ref)
```


## MCMC | constant M

```{r MCMC constant}
load("~/pCloudDrive/ICCAT/runs/a4a/MCMC/final_constM.RData")
plot(FLStocks(MLE = a4a.stk9_sim, MCMC = aa.stk + fit20))
```

## MCMC | constant M discards included

```{r MCMC constant discards}
load("~/pCloudDrive/ICCAT/runs/a4a/MCMC/final_constM_disc.RData")
plot(FLStocks(MLE = a4a.stk9_disc_sim, MCMC = aa.stk + fit40))
```