---
title: '**Assessing Mediterranean Swordfish Using the Assessment for All (a4a) Model \vspace{0.1cm}**'
documentclass: article
author: "\\small D. Mantopoulou-Palouka$^1$, G. Tserpes$^2$, A. Kimoto$^3$, E. Jardim$^4$"
date: ''
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
mainfont: Caladea
fontsize: 10pt
geometry: a4paper, left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm
header-includes:
- \usepackage{titling}
- \pretitle{\begin{center}\large\bfseries}
- \posttitle{\end{center}}
- \usepackage{soul,color,colortbl,fancyhdr}
- \usepackage{caption}
- \usepackage[singlelinecheck=false]{caption}
- \usepackage[document]{ragged2e}
---

\footnotetext[1]{Joint Research Center. Corresponding author: Danai.MANTOPOULOU-PALOUKA@ec.europa.eu}
\footnotetext[2]{Hellenic Center for Marine Research, Heraklion, Greece}
\footnotetext[3]{ International Commission for the Conservation of Atlantic Tunas, Madrid Spain}
\footnotetext[4]{Portuguese Institute for Sea and Atmosphere, Lisbon, Portugal}

\pagestyle{fancy}
\fancypagestyle{plain}{\pagestyle{fancy}}
\fancyhead{}
\fancyhead[L]{SCRS/2026/169}
\captionsetup[table]{labelformat=empty}


```{r setup, include=FALSE}
## knit from the project root so that "Robj/..." paths resolve
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.width = 7, fig.height = 4.5, dpi = 300,
  cache = TRUE, cache.lazy = FALSE
)
```

```{r libs, cache=FALSE}
suppressMessages({
  library(FLCore)
  library(FLa4a)
  library(FLBRP)
  library(ggplotFL)
  library(knitr)
})
theme_set(theme_bw(base_size = 10))
```

```{r load-data}
## 2020 assessment inputs (ages 0-5, 1985-2018, plus group 5)
swo.stk  <- readRDS("Robj/swo_stk.rds")
swo.idx  <- readRDS("Robj/swo_bio_idx_std.rds")
swo.stk.l <- readRDS("Robj/swo_stk_loren.rds") # lorenzen
source("analysis/plot_functions.R")
```

## Input data


```{r caaPlot, fig.cap = "Catch at age (ages 0-9, 1972-2024)"}
ggplot(as.data.frame(catch.n(swo.stk)), 
       aes(x = age, 
           y = data, 
           color = as.factor(floor(as.numeric(as.character(year)) / 10) * 10), 
           group = as.factor(year))) +
  geom_line() +
  labs(color = "Decade")+
  scale_color_viridis_d(option = "turbo")+
  geom_vline(xintercept = c(1,2,3), linetype = 'dotted')
```

## Candidate runs

### Long time series 

A bit unstable especially age 2 and 3 residuals in the beginning of the time sereis


```{r, echo = TRUE}
stk <- window(swo.stk, start = 1978)
stk <- replaceZeros(stk)
range(stk)[c('minfbar','maxfbar')] <- c(1,4) 
idx <- swo.idx[-2]

#srmod <- ~s(year, k =10, by = as.numeric(year>1987))
fmod3 <- ~s(year, k = 25) + s(age, k=8) + ti(year, age, k = c(6,3)) + s(year, k =10, by = as.numeric(age==3))
qmod <- list(~1, ~s(year, k = 3), ~1, ~s(year, k = 3),~1)
#fit <- sca(stk, idx, fmodel = fmod3, qmodel = qmod, srmodel = srmod)
fit <- sca(stk, idx, fmodel = fmod3, qmodel = qmod) 

a4a.stk <- stk + simulate(fit, nsim = 1000)
plot(a4a.stk, stk)+ylim(0.01, NA)
res <- residuals(fit, stk, idx)
plot(res)
plot_status(stk,fit)
```