---
title: "Assessing Mediterranean Swordfish Using the Assessment for All (a4a) Model"
shorttitle: "Mediterranean swordfish a4a exploratory runs"
author: "D. Mantopoulou-Palouka, G. Tserpes, A. Kimoto, E. Jardim"
institute: ""
date: "SCRS/2026/170"
output:
  beamer_presentation:
    latex_engine: xelatex
    theme: "Copenhagen"
    colortheme: "seahorse"
    slide_level: 2
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{TeX Gyre Termes}
  - \usepackage{booktabs}
  - \usepackage{longtable}
---

```{r setup, include=FALSE}
## Knit from project root so paths resolve correctly
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.width = 7, fig.height = 4.3, dpi = 300,
  cache = TRUE, cache.lazy = FALSE
)
```

```{r libs, cache=FALSE}
suppressMessages({
  library(FLCore)
  library(FLa4a)
  library(FLBRP)
  library(ggplotFL)
  library(knitr)
  library(kableExtra)
  library(gridExtra)
})
theme_set(theme_bw(base_size = 9))
source("analysis/plot_functions.R")
## Helper for custom retro with dynamic F/SR submodels
retro_runs <- function(stk, idxs, retro = 5, kfrac = 0.3, k_f = 20, k_sr = 20, qmod) {
  lst0 <- split(0:retro, 0:retro)
  lst0 <- lapply(lst0, function(x) {
    yr <- range(stk)["maxyear"] - x
    stk_sub <- window(stk, end = yr)
    idx_sub <- FLIndices(window(idxs, end = yr))
    KY <- unname(k_f - floor(x * kfrac))
    KZ <- unname(k_sr - floor(x * kfrac))
    fmod_sub <- as.formula(substitute(~ s(year, k = KY) + s(age, k = 8) + ti(year, age, k = c(6, 5)), list(KY = KY)))
    srmod_sub <- as.formula(substitute(~ s(year, k = KZ), list(KZ = KZ)))
    fit_sub <- sca(stk_sub, idx_sub, fmodel = fmod_sub, srmodel = srmod_sub, qmodel = qmod)
    stk_sub + fit_sub
  })
  FLStocks(lst0)
}
```

```{r load-data}
swo.stk   <- readRDS("Robj/swo_stk.rds")
swo.stk.l <- readRDS("Robj/swo_stk_loren.rds")
swo.idx   <- readRDS("Robj/swo_bio_idx_std.rds")

## Use standardized longline indices (excluding index 2)
idx_use <- swo.idx[-2]
for (i in seq_along(idx_use)) {
  range(idx_use[[i]], c("min", "max")) <- c(1, 4)
  units(index(idx_use[[i]])) <- "t"
}
```

## Background

- **Context:** Following the update of the 2020 baseline $\alpha4\alpha$ assessment through 2024, structural enhancements were investigated to address historical data coverage and biological realism. The new catch at age matrix support using a plus group at age 9.
- **Key Objectives:**
  - **Temporal Scope:** Evaluate sensitivity to the time-series start year by comparing the truncated period (**1985–2024**) against the extended historical series (**1978–2024**).
  - **Natural Mortality ($M$):** Assess the impact of assuming a **Constant $M$** versus an age-dependent **Lorenzen $M$** vector.
  - **Fishing Mortality Dynamics:** Incorporate a tensor product interaction (`ti(year, age)`) to capture time-varying selectivity across years and age classes.

## Input Data and Assumptions
                                                       
Catch at age 0-9 with plus group set at 9

```{r inputData1}
    p1 <- ggplot(as.data.frame(catch.n(swo.stk)), 
                 aes(x = as.numeric(as.character(age)), 
                     y = data, 
                     color = as.factor(floor(as.numeric(as.character(year)) / 10) * 10),              
                     group = as.factor(year))) +
      geom_line(alpha = 0.85) +
      geom_vline(xintercept = c(1, 2, 3), linetype = "dotted") +
      scale_color_viridis_d(option = "turbo") +
      labs(title = "Catch-at-Age", x = "Age", y = "Numbers", color = "Decade") +
      theme_bw(base_size = 9) +
      theme(legend.position = "bottom")+
      scale_x_continuous(breaks = c(0:9))
  
    
    mat_df <- as.data.frame(mat(swo.stk)[, 1])
    p2 <- ggplot(mat_df, aes(x = as.numeric(as.character(age)), y = data)) +
      geom_line(color = "#2b83ba", linewidth = 0.9) +
      geom_point(color = "#2b83ba", size = 1.6) +
      labs(title = "Maturity Ogive", x = "Age", y = "Proportion Mature") +
      theme_bw(base_size = 8.5)+
      scale_x_continuous(breaks = c(0:9))
  
    
    m_df <- rbind(
      data.frame(age = as.numeric(as.character(as.data.frame(m(swo.stk)[, 1])$age)),
                 M   = as.data.frame(m(swo.stk)[, 1])$data,
                 Type = "Constant M"),
      data.frame(age = as.numeric(as.character(as.data.frame(m(swo.stk.l)[, 1])$age)),
                 M   = as.data.frame(m(swo.stk.l)[, 1])$data,
                 Type = "Lorenzen M")
    )
  
    p3 <- ggplot(m_df, aes(x = age, y = M, color = Type)) +
      geom_line(linewidth = 0.9) +
      geom_point(size = 1.6) +
      scale_color_manual(values = c("Constant M" = "#d7191c", "Lorenzen M" = "#fdae61")) +            
      labs(title = "Natural Mortality (M)", x = "Age", y = "M", color = NULL) +
      theme_bw(base_size = 8.5) +
      theme(legend.position = "bottom")+
      scale_x_continuous(breaks = c(0:9))
  
    ## Combined Layout
    (p1 | (p2 / p3)) + plot_layout(widths = c(1.3, 1))

```

<!-- ## Input Data and Assumptions -->

```{r inputData2, fig.height = 4.1}

```

## Summary of Candidate & Sensitivity Runs

```{r runs-table}
runs_df <- data.frame(
  Run = c("Run 1 ", "Run 2 ", 
          "Sensitivity 1 ", "Sensitivity 2 "),
  Period = c("1985–2024", "1985–2024", "1978–2024", "1985–2024"),
  `M Assumption` = c("Constant M", "Lorenzen M", "Constant M", "Constant M"),
  `Fbar Range` = c("Ages 1–4", "Ages 1–4", "Ages 1–4", "Ages 1–4"),
  `Indices` = c("5 indices", "5 indices", "5 indices", "5 indices + SI_GN + LI_SUR"),
  check.names = FALSE
)
kable(runs_df, row.names = FALSE, booktabs = TRUE)
```

## Model Specifications & Sub-models

```{r submodels-table}
submods_df <- data.frame(
  Component = c("F sub-model (1985 runs)", 
                # "F sub-model (1978 runs)", 
                "SR sub-model", "Q sub-model (GR, SP, SI)", 
                "Q sub-model (LI, MO)"),
  Specification = c("~ s(year, k = 20) + s(age, k = 8) + ti(year, age, k = c(6, 5))",
                    # "~ s(year, k = 25) + s(age, k = 8) + ti(year, age, k = c(6, 5))",
                    "~ s(year, k = 20)",
                    "~ 1 (Constant catchability)",
                    "~ s(year, k = 3) (Time-varying catchability)"),
  check.names = FALSE
)
kable(submods_df, row.names = FALSE, booktabs = TRUE)
```

```{r fit-1985-runs}
## Run 1: 1985 Constant M
stk85_c <- replaceZeros(window(swo.stk, start = 1985))
range(stk85_c)[c('minfbar', 'maxfbar')] <- c(1, 4)
fmod_85 <- ~ s(year, k = 20) + s(age, k = 8) + ti(year, age, k = c(6, 5))
srmod   <- ~ s(year, k = 20)
qmod    <- list(~1, ~s(year, k = 3), ~1, ~s(year, k = 3), ~1)

fit85_c <- sca(stk85_c, idx_use, fmodel = fmod_85, srmodel = srmod, qmodel = qmod)
a4a85_c <- stk85_c + simulate(fit85_c, nsim = 1000)

save(stk85_c,fit85_c, idx_use, file = "Robj/input4mcmc.Rdata")

## Run 2: 1985 Lorenzen M
stk85_l <- replaceZeros(window(swo.stk.l, start = 1985))
range(stk85_l)[c('minfbar', 'maxfbar')] <- c(1, 4)
fit85_l <- sca(stk85_l, idx_use, fmodel = fmod_85, srmodel = srmod, qmodel = qmod)
a4a85_l <- stk85_l + simulate(fit85_l, nsim = 1000)
```

```{r fit-1978-runs}
## Sensitivity 1: 1978 Constant M
stk78_c <- replaceZeros(window(swo.stk, start = 1978))
range(stk78_c)[c('minfbar', 'maxfbar')] <- c(1, 4)
fmod_78 <- ~ s(year, k = 25) + s(age, k = 8) + ti(year, age, k = c(6, 5))

fit78_c <- sca(stk78_c, idx_use, fmodel = fmod_78, srmodel = srmod, qmodel = qmod)
a4a78_c <- stk78_c + simulate(fit78_c, nsim = 1000)

## Sensitivity 2: 1985 Constant M - extra indices
stk_s2 <- replaceZeros(window(swo.stk.l, start = 1985))
idx_s2 <- readRDS("Robj/swo_bio_idx_std_ext.rds")
range(stk_s2)[c('minfbar', 'maxfbar')] <- c(1, 4)
qmod_s2    <- list(~1, ~s(year, k = 5), ~s(year, k = 3), ~1, ~s(year, k = 3), ~1, ~1)
fit_s2 <- sca(stk_s2, idx_s2, fmodel = fmod_78, srmodel = srmod, qmodel = qmod_s2)
a4a_s2 <- stk_s2 + simulate(fit_s2, nsim = 1000)
```

## Trajectory Comparison: 1985 Series (Constant vs. Lorenzen M)

```{r plot-1985-comp, fig.height=4.2}
plot(FLStocks(`1985 Constant M` = a4a85_c, `1985 Lorenzen M` = a4a85_l))+ylim(0.01,NA)
```

<!-- ## Trajectory Comparison: 1978 Series (Constant vs. Lorenzen M) -->

```{r plot-1978-comp, fig.height=4.2}
# plot(FLStocks(`1978 Constant M` = a4a78_c, `1978 Lorenzen M` = a4a78_l))
```

## Candidate Runs vs. Observed Catch

```{r plot-all-comp, fig.height=4.2}
# plot(FLStocks(
#   `85 Const M`  = a4a85_c,
#   `85 Loren M`  = a4a85_l,
#   `78 Const M`  = a4a78_c,
#   `78 Loren M`  = a4a78_l
# ))
plot(FLStocks(
  `85 Const M`  = a4a85_c,
  `85 Loren M`  = a4a85_l,
  `Catch` = stk85_c
)) + ylim(0.01,NA)
```

# Diagnostics - Constant M

## Catch-at-Age Standardized Residuals

```{r diag-caa-res, fig.height=5.2}
res85_c <- residuals(fit85_c, stk85_c, idx_use)
# res_df  <- as.data.frame(res85_c)
# ggplot(subset(res_df, qname == "catch.n"), aes(year, data)) +
#   geom_point(alpha = 0.6) +
#   geom_smooth(se = FALSE, color = "blue", size = 0.7) +
#   facet_wrap(~age, ncol = 3, scales = "free_y") +
#   labs(x = "Year", y = "Standardized Residuals")

    res_df <- as.data.frame(residuals(fit85_c, stk85_c, idx_use))
  
    ggplot(subset(res_df, qname == "catch.n"), aes(x = year, y = data)) +
      geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
      geom_segment(aes(xend = year, yend = 0, color = data > 0), linewidth = 0.5) +
      geom_point(aes(color = data > 0), size = 1.5) +
      scale_color_manual(values = c("TRUE" = "#2c7bb6", "FALSE" = "#d7191c"), 
                         labels = c("TRUE" = "Positive", "FALSE" = "Negative")) +
      facet_wrap(~ age, ncol = 3) +
      labs(x = "Year", y = "Standardized Residuals", color = "Residual Sign") +
      theme_bw() +
      theme(legend.position = "bottom") + ylim(-3,3)

```

## Standardized CPUE Index Residuals
```{r diag-idx-res, fig.height=5.2}
# ggplot(subset(res_df, !qname %in% c("catch.n", "catch")), aes(year, data)) +
#   geom_point(alpha = 0.7) +
#   geom_smooth(se = FALSE, color = "darkred", size = 0.7) +
#   facet_wrap(~qname, ncol = 3) +
#   labs(x = "Year", y = "Standardized Residuals")

    ggplot(subset(res_df, !qname %in% c("catch.n", "catch")), aes(x = year, y = data)) +              
      geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
      geom_segment(aes(xend = year, yend = 0, color = data > 0), linewidth = 0.5) +
      geom_point(aes(color = data > 0), size = 1.5) +
      scale_color_manual(values = c("TRUE" = "#2c7bb6", "FALSE" = "#d7191c"), 
                         labels = c("TRUE" = "Positive", "FALSE" = "Negative")) +
      facet_wrap(~ qname, ncol = 3) +
      labs(x = "Year", y = "Standardized Residuals", color = "Residual Sign") +
      theme_bw() +
      theme(legend.position = "bottom")

```

## Catch-at-Age Bubble Plot

```{r diag-bubbles, fig.height=5}
bubbles(res85_c)
```

## Aggregated catch diagnostics

```{r aggr-catch, fig.height=4}
cd <- computeCatchDiagnostics(fit85_c, stk85_c)
obs_df <- as.data.frame(cd[["obs"]]); obs_df <- obs_df[, c("year", "data")]
est_df <- as.data.frame(cd[["est"]]); est_df <- est_df[, c("year", "data")]
    ## 1. Top-Left: Observation Error (oe)
    p1 <- plot_catch_err(cd[["oe"]], "Observation Error (oe)")
  
    ## 2. Top-Right: Estimation Error (ee)
    p2 <- plot_catch_err(cd[["ee"]], "Estimation Error (ee)")
  
    ## 3. Bottom-Left: Combined Error (oee)
    p3 <- plot_catch_err(cd[["oee"]], "Prediction Error (oee)")
  
    ## 4. Bottom-Right: Standardized Catch Residuals (resstd)
    res_df <- as.data.frame(cd[["resstd"]])
    p4 <- ggplot(res_df, aes(x = year, y = data)) +
      geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
      geom_segment(aes(xend = year, yend = 0, color = data > 0), linewidth = 0.6) +
      geom_point(aes(color = data > 0), size = 1.8) +
      geom_smooth(se = FALSE, color = "black", linewidth = 0.6) +
      scale_color_manual(values = c("TRUE" = "#2b83ba", "FALSE" = "#d7191c")) +
      labs(title = "Standardized Catch Residuals (resstd)", x = "Year", y = "Residuals") +            
      theme_bw(base_size = 9) +
      theme(legend.position = "none")
  
    ## Combine into a 2x2 layout
    (p1 | p2) / (p3 | p4)
```

## Observed vs. Fitted Stock - Catch (85 Const M)

```{r diag-obs-fit, fig.height=5.5}
plot(fit85_c, stk85_c) 
```

## Retrospective Analysis (Peels = 5)

```{r retro-85-c, fig.height=4.1}
retro_85c <- retro_runs(stk85_c, idx_use, retro = 5, k_f = 20, k_sr = 20, qmod = qmod)
mohn_ssb <- icesAdvice::mohn(mohnMatrix(retro_85c, ssb))
mohn_f   <- icesAdvice::mohn(mohnMatrix(retro_85c, fbar))
plot(retro_85c) + theme(legend.position = "none")+
      labs(
        subtitle = sprintf("Mohn's rho: SSB = %.3f  |  Fbar = %.3f", mohn_ssb, mohn_f)
      )
```

## Stock status

```{r stockS-85-c, fig.heigt = 5.5}
plot_status(stk85_c, fit85_c, a4a85_c) + labs(subtitle = "Beverton-Holt SRR | b fixed at 100")
```

<!-- ================================================================================================== -->

# Diagnostics - Lorenzen M

## Catch-at-Age Standardized Residuals
```{r diag-caa-res-l, fig.height=5.2}
res85_l <- residuals(fit85_l, stk85_l, idx_use)

    res_df <- as.data.frame(residuals(fit85_l, stk85_l, idx_use))
  
    ggplot(subset(res_df, qname == "catch.n"), aes(x = year, y = data)) +
      geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
      geom_segment(aes(xend = year, yend = 0, color = data > 0), linewidth = 0.5) +
      geom_point(aes(color = data > 0), size = 1.5) +
      scale_color_manual(values = c("TRUE" = "#2c7bb6", "FALSE" = "#d7191c"), 
                         labels = c("TRUE" = "Positive", "FALSE" = "Negative")) +
      facet_wrap(~ age, ncol = 3) +
      labs(x = "Year", y = "Standardized Residuals", color = "Residual Sign") +
      theme_bw() +
      theme(legend.position = "bottom") + ylim(-3,3)

```

## Standardized CPUE Index Residuals

```{r diag-idx-res-l, fig.height=5.2}

    ggplot(subset(res_df, !qname %in% c("catch.n", "catch")), aes(x = year, y = data)) +              
      geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
      geom_segment(aes(xend = year, yend = 0, color = data > 0), linewidth = 0.5) +
      geom_point(aes(color = data > 0), size = 1.5) +
      scale_color_manual(values = c("TRUE" = "#2c7bb6", "FALSE" = "#d7191c"), 
                         labels = c("TRUE" = "Positive", "FALSE" = "Negative")) +
      facet_wrap(~ qname, ncol = 3) +
      labs(x = "Year", y = "Standardized Residuals", color = "Residual Sign") +
      theme_bw() +
      theme(legend.position = "bottom")

```


## Catch-at-Age Bubble Plot

```{r diag-bubbles-l, fig.height=5}
bubbles(res85_l)
```

## Aggregated catch diagnostics

```{r aggr-catch-l, fig.height=4}
cd <- computeCatchDiagnostics(fit85_l, stk85_l)
obs_df <- as.data.frame(cd[["obs"]]); obs_df <- obs_df[, c("year", "data")]
est_df <- as.data.frame(cd[["est"]]); est_df <- est_df[, c("year", "data")]
    ## 1. Top-Left: Observation Error (oe)
    p1 <- plot_catch_err(cd[["oe"]], "Observation Error (oe)")
  
    ## 2. Top-Right: Estimation Error (ee)
    p2 <- plot_catch_err(cd[["ee"]], "Estimation Error (ee)")
  
    ## 3. Bottom-Left: Combined Error (oee)
    p3 <- plot_catch_err(cd[["oee"]], "Prediction Error (oee)")
  
    ## 4. Bottom-Right: Standardized Catch Residuals (resstd)
    res_df <- as.data.frame(cd[["resstd"]])
    p4 <- ggplot(res_df, aes(x = year, y = data)) +
      geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
      geom_segment(aes(xend = year, yend = 0, color = data > 0), linewidth = 0.6) +
      geom_point(aes(color = data > 0), size = 1.8) +
      geom_smooth(se = FALSE, color = "black", linewidth = 0.6) +
      scale_color_manual(values = c("TRUE" = "#2b83ba", "FALSE" = "#d7191c")) +
      labs(title = "Standardized Catch Residuals (resstd)", x = "Year", y = "Residuals") +            
      theme_bw(base_size = 9) +
      theme(legend.position = "none")
  
    ## Combine into a 2x2 layout
    (p1 | p2) / (p3 | p4)
```

## Observed vs. Fitted Stock - Catch

```{r diag-obs-fit-l, fig.height=5.5}
plot(fit85_l, stk85_l)
```

## Retrospective Analysis (Peels = 5)

```{r retro-85-l, fig.height=4.1}
retro_85l <- retro_runs(stk85_l, idx_use, retro = 5, k_f = 20, k_sr = 20, qmod = qmod)
mohn_ssb <- icesAdvice::mohn(mohnMatrix(retro_85l, ssb))
mohn_f   <- icesAdvice::mohn(mohnMatrix(retro_85l, fbar))
plot(retro_85l) + theme(legend.position = "none")+
      labs(
        subtitle = sprintf("Mohn's rho: SSB = %.3f  |  Fbar = %.3f", mohn_ssb, mohn_f)
      )
```

## Stock status

```{r stockS-85-l, fig.heigt = 5.5}
plot_status(stk85_l, fit85_l, a4a85_l)+ labs(subtitle = "Beverton-Holt SRR | b fixed at 100")
```


# Candidate runs comparison

## Fishing mortality 3D plots

```{r harvest3D, fig.height=4}
aa4plot_c <- stk85_c + fit85_c
aa4plot_l <- stk85_l + fit85_l
w1 <- wireframe(data ~ age + year, data = as.data.frame(harvest(aa4plot_c)), drape = TRUE,
          main="Fishing mortality", screen = list(x = -90, y=-45),
          col.regions = colorRampPalette(c("blue", "yellow"))(100))
w2 <- wireframe(data ~ age + year, data = as.data.frame(harvest(aa4plot_l)), drape = TRUE,
          main="Fishing mortality", screen = list(x = -90, y=-45),
          col.regions = colorRampPalette(c("blue", "yellow"))(100))
grid.arrange(w1,w2,ncol = 2)
```

## Stock Status & Reference Points 
### Candidate Runs Comparison

```{r refpts-calc}
calc_refpts <- function(stk, fit) {
  aa_stk <- stk + fit
  landings.n(aa_stk) <- catch.n(aa_stk)
  landings.wt(aa_stk) <- catch.wt(aa_stk)
  landings(aa_stk)   <- catch(aa_stk)
  discards(aa_stk)   <- 0
  discards.n(aa_stk) <- 0
  discards.wt(aa_stk) <- 0
  srr <- fmle(as.FLSR(aa_stk, model = "bevholt"), control = list(trace = 0), fixed = list(b = 100))
  brp_obj <- brp(FLBRP(aa_stk, sr = srr))
  refpts(brp_obj)
}

rp_85c <- calc_refpts(stk85_c, fit85_c)
rp_85l <- calc_refpts(stk85_l, fit85_l)
aa4ref_c <- stk85_c+fit85_c
aa4ref_l <- stk85_l+fit85_l

ref_table <- data.frame(
  Metric = c("FMSY", "BMSY (t)", "F2024", "B2024 (t)", "F2024 / FMSY", "B2024 / BMSY","MSY"),
  `1985 Constant M` = round(c(
    as.numeric(rp_85c['msy', 'harvest']),
    as.numeric(rp_85c['msy', 'biomass']),
    as.numeric(fbar(aa4ref_c)[, dim(fbar(aa4ref_c))[2]]),
    as.numeric(stock(aa4ref_c)[, dim(stock(aa4ref_c))[2]]),
    as.numeric(fbar(aa4ref_c)[, dim(fbar(aa4ref_c))[2]] / rp_85c['msy', 'harvest']),
    as.numeric(stock(aa4ref_c)[, dim(stock(aa4ref_c))[2]] / rp_85c['msy', 'biomass']),
    as.numeric(rp_85c['msy', 'yield'])
  ), 3),
  `1985 Lorenzen M` = round(c(
    as.numeric(rp_85c['msy', 'harvest']),
    as.numeric(rp_85c['msy', 'biomass']),
    as.numeric(fbar(aa4ref_l)[, dim(fbar(aa4ref_l))[2]]),
    as.numeric(stock(aa4ref_l)[, dim(stock(aa4ref_l))[2]]),
    as.numeric(fbar(aa4ref_l)[, dim(fbar(aa4ref_l))[2]] / rp_85l['msy', 'harvest']),
    as.numeric(stock(aa4ref_l)[, dim(stock(aa4ref_l))[2]] / rp_85l['msy', 'biomass']),
    as.numeric(rp_85c['msy', 'yield'])
  ), 3),
  check.names = FALSE
)

kable(ref_table, row.names = FALSE, booktabs = TRUE)
```

## Stock Status Trajectories (F/FMSY vs. B/BMSY)

```{r kobe-85-c, fig.heigt = 5.5}
p1 <- plot_kobe(a4a85_c, rp_85c, use_ssb = FALSE, title = "Kobe Plot: 1985 Constant M")   
p2 <- plot_kobe(a4a85_l, rp_85l, use_ssb = FALSE, title = "Kobe Plot: 1985 Lorenzen M")  
p1 + p2
```

## Sensitivity runs

```{r sens-plots, fig.height = 5.5}
p1 <- plot(FLStocks(Sensitivity1 = a4a78_c, Sensitivity2 = a4a_s2))+theme(legend.position = "top")
p2 <- plot(FLStocks(Sensitivity1 = a4a78_c, Sensitivity2 = a4a_s2,
Candidate1 = a4a85_c, Candidate2 = a4a85_l))+theme(legend.position = "top")
p1 + p2
```

## Sensitivity runs - Stock status

```{r sesns-plots-status}
p1 <- plot_status(stk = stk78_c,fit = fit78_c, a4a.stk = a4a78_c)
p2 <- plot_status(stk = stk_s2,fit = fit_s2, a4a.stk = a4a_s2)
p1+p2
```

<!-- ## Conclusions & Recommendations for SCRS -->

<!-- ## Conclusions -->

<!-- - **Model Stability:** Incorporating the tensor product interaction `ti(year, age)` resolves residual patterns in catch-at-age compared to additive age-year structures.
- **Natural Mortality:** Lorenzen age-varying $M$ scales absolute abundance estimates higher for younger age classes while preserving relative trends and fishing mortality trajectories.
- **Time-Series Scope:** Results starting in **1985** display robust diagnostic performance and consistency with standardized longline abundance indices. -->

<!-- ## Thank you -->

## Additional run without Sicilian LL 

Proposal of the group: Exclude Sicilian LL - use only four indices

```{r excl-SILL}
## Proposal of the group: Exclude Sicilian LL - use only four indices
idx_gp <- idx_use[-5]
fmod_85 <- ~ s(year, k = 20) + s(age, k = 8) + ti(year, age, k = c(6, 5))
srmod   <- ~ s(year, k = 20)
qmod_gp    <- list(~1, ~s(year, k = 3), ~1, ~s(year, k = 3))
fit_gp <- sca(stk85_c, idx_gp, fmodel = fmod_85, srmodel = srmod, qmodel = qmod_gp)
a4a_gp <- stk85_c + simulate(fit_gp, nsim = 1000)
```

```{r plt-cmp-bc-gp, fig.height = 5}
plot(FLStocks(`4 indices` = a4a_gp, `Constant M` = a4a85_c))
```

## Catch-at-Age Standardized Residuals - Excluding Sicilian LL 
```{r diag-caa-res-gp, fig.height=5.2}
res85_l <- residuals(fit_gp, stk85_c, idx_gp)

    res_df <- as.data.frame(residuals(fit_gp, stk85_c, idx_gp))
  
    ggplot(subset(res_df, qname == "catch.n"), aes(x = year, y = data)) +
      geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
      geom_segment(aes(xend = year, yend = 0, color = data > 0), linewidth = 0.5) +
      geom_point(aes(color = data > 0), size = 1.5) +
      scale_color_manual(values = c("TRUE" = "#2c7bb6", "FALSE" = "#d7191c"), 
                         labels = c("TRUE" = "Positive", "FALSE" = "Negative")) +
      facet_wrap(~ age, ncol = 3) +
      labs(x = "Year", y = "Standardized Residuals", color = "Residual Sign") +
      theme_bw() +
      theme(legend.position = "bottom") + ylim(-3,3)

```

## Standardized CPUE Index Residuals - Excluding Sicilian LL

```{r diag-idx-res-gp, fig.height=5.2}

    ggplot(subset(res_df, !qname %in% c("catch.n", "catch")), aes(x = year, y = data)) +              
      geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
      geom_segment(aes(xend = year, yend = 0, color = data > 0), linewidth = 0.5) +
      geom_point(aes(color = data > 0), size = 1.5) +
      scale_color_manual(values = c("TRUE" = "#2c7bb6", "FALSE" = "#d7191c"), 
                         labels = c("TRUE" = "Positive", "FALSE" = "Negative")) +
      facet_wrap(~ qname, ncol = 3) +
      labs(x = "Year", y = "Standardized Residuals", color = "Residual Sign") +
      theme_bw() +
      theme(legend.position = "bottom")

```