---
title: "Continuity model runs for the Mediterranean swordfish stock assessment using
  the a4a initiative"
shorttitle: "Mediterranean swordfish a4a continuity runs"
author: "D. Mantopoulou-Palouka, G. Tserpes, A. Kimoto, E. Jardim"
institute: ""
date: "SCRS/2026/169"
output:
  beamer_presentation:
    latex_engine: xelatex
    theme: "Copenhagen"
    colortheme: "seahorse"
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{TeX Gyre Termes}
---


```{r setup, include=FALSE}
## knit from the project root so that "Robj/..." paths resolve
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.width = 7, fig.height = 4.5, dpi = 300,
  cache = TRUE, cache.lazy = FALSE
)
```


```{r libs, cache=FALSE}
suppressMessages({
  library(FLCore)
  library(FLa4a)
  library(FLBRP)
  library(ggplotFL)
  library(knitr)
})
theme_set(theme_bw(base_size = 10))
```

```{r load-data}
## 2020 assessment inputs (ages 0-5, 1985-2018, plus group 5)
stk_2020 <- readRDS("Robj/swo_stk_2020.rds")
idx_2020 <- readRDS("Robj/swo_bio_idx_2020.rds")
swo.stk  <- readRDS("Robj/swo_stk.rds")
swo.idx  <- readRDS("Robj/swo_bio_idx_std.rds")
```


## Objectives

To evaluate current stock status by integrating recent fisheries data into the established 2020 α4α model configuration. 

- Fundamental biological assumptions and model structures remain the same. 

- The update focuses on incorporating standardized relative abundance indices and catch-at-age data for 2020–2024, followed by a diagnostic review and revised stock status estimates.

## Available catch-at-age data

Catch-at-age (CAA) matrices for ages 0-9 spanning `r range(swo.stk)["minyear"]`-`r range(swo.stk)["maxyear"]` were provided by the ICCAT Secretariat. 

```{r caa-comp, fig.height=4, fig.cap="", results='asis'}
ggplot(as.data.frame(FLQuants(caa2020 = catch.n(stk_2020), 
                              caa2026 = catch.n(swo.stk)))) +
  geom_line(aes(x = year, y = data, group = qname, color = qname))+
  facet_wrap(~age, ncol = 3)
```

## Input catch-at-age data
For continuity with 2020, the fitted model truncates the series to start in 1985 and collapses ages into a plus group at age 5, with the reference fishing mortality ($\bar{F}$) averaged over ages 2-4.

```{r caa-comp2, fig.height=3, fig.cap="", echo=FALSE, results='asis', message = FALSE, warning=FALSE, comment=NA}
stk <- window(swo.stk, start = 1985) # truncate early years as in 2020
stk <- replaceZeros(stk)
range(stk)[c('minfbar','maxfbar')] <- c(2,4) 
stk <- setPlusGroup(stk, 5) # same plus group
#range(stk)[c('minfbar','maxfbar')] <- c(2,4) 
ggplot(as.data.frame(FLQuants(caa2020 = catch.n(stk_2020), 
                              caa2026 = catch.n(stk)))) +
  geom_line(aes(x = year, y = data, group = qname, color = qname))+
  facet_wrap(~age, ncol = 3, scales = 'free')
```

## CPUE series
Used in the 2020 assessment: 

```{r idx-tab, table.placement = 'H'}
library(kableExtra)

idx_labels <- c(
  GR_LL       = "Greek longliners",
  LI_LL       = "Ligurian longliners",
  SP_LL       = "Spanish longliners",
  MO_LL       = "Moroccan longliners",
  SI_LL       = "Sicilian longliners"
)
idx_t <- swo.idx[-2]
idx_summary <- do.call(rbind, lapply(names(idx_t), function(nm){
  x <- idx_t[[nm]]
  data.frame(
    Index        = idx_labels[[nm]],
    Series       = nm,
    `First year` = range(x)["minyear"],
    `Last year`  = range(x)["maxyear"],
    `Age range`  = paste(range(x)["min"], range(x)["max"], sep = "-"),
    check.names  = FALSE,
    row.names    = NULL)
}))
kable(idx_summary, row.names = FALSE)
      #caption = "Table 1: Standardized CPUE series used as relative abundance indices in the 2026 updated a4a assessment.") %>% 
  #kable_styling(latex_options = "hold_position")
```

## Statistical analysis

Four runs are reported: 

- The 2020 a4a model as a baseline 

- The updated run fitted to data through 2024

- A sensitivity run that truncate the updated data to the 2018, terminal year of the 2020 assessment

- A sensitivity run that truncate the updated data to 2018, using the 2020 standardized indices. 


## Fishing mortality & recruitment sub-models

Under the `a4a` framework, fishing mortality, recruitment and catchability are specified
as separable sub-models of smooth functions of age and year. 


```{r mods}
mods <- data.frame(
  Run = c("2020 baseline", "2026 update", "Sensitivity 1 (up to 2018)", "Sensitivity 2 (up to 2018 - old indices)"),
  `F submodel` = c("s(year, k=17) + s(age, k=3)",
                   "s(year, k=19) + s(age, k=3)",
                   "s(year, k=17) + s(age, k=3)",
                   "s(year, k=17) + s(age, k=3)"),
  `R submodel` = c("s(year, k=15)", "s(year, k=17)", "s(year, k=15)", "s(year, k=15)"),
  check.names = FALSE)
kable(mods, row.names = FALSE)
      #caption = "Table 2: Fishing-mortality and recruitment submodels for the reported a4a runs")  %>% 
  #kable_styling(latex_options = "hold_position")
```

## Catchability sub-model

```{r qmods}
qmod_table <- data.frame(
  Index_Code = c("GR_LL", "LI_LL", "SP_LL", "MO_LL", "SI_LL"),
  Fleet = c("Greek Longline", 
            "Ligurian Longline", 
            "Spanish Longline", 
            "Moroccan Longline", 
            "Sicilian Longline"),
  Submodel = c("~ 1", 
              "~ s(year, k = 3)", 
              "~ 1", 
              "~ s(year, k = 3)", 
              "~ 1")
)

# Render the table
kable(qmod_table, 
      col.names = c("Index", "Fleet", "qmodel Formula"),
      align = "llcl")  %>% 
  kable_styling(latex_options = "hold_position")
```

## Model results - 2020 vs 2026 update 

```{r fit-2020}
# Dropping the Ligurian surface (as we did in 2020)
idx_2020 <- idx_2020[-6]
# NOTE: final a4a had discards included
fmod2020 <- ~s(year, k = 17) + s(age, k = 3)
srmod2020 <- ~s(year, k = 15)
# qmod2020 <- list(~1,~1, ~s(year, k = 3), ~s(year, k = 3),~1,~s(year, k = 9))
qmod2020 <- list(~1,~1, ~s(year, k = 3), ~s(year, k = 3),~1)
fit2020 <- sca(stk_2020, idx_2020, fmodel = fmod2020, srmodel = srmod2020, qmodel = qmod2020)
a4a.2020 <- stk_2020 + simulate(fit2020, nsim = 1000)
```

```{r fit-update}
# Also dropping the Ligurian from here
idx <- swo.idx[-2]

fmod2 <- ~s(year, k = 20) + s(age, k = 3) # added ks - keep ratio
srmod1 <- ~s(year, k = 17) # added ks
# qmod <- list(~1,~s(year, k = 9), ~s(year, k = 3), ~1, ~s(year, k = 3),~1)
qmod <- list(~1, ~s(year, k = 3), ~1, ~s(year, k = 3),~1)
fit <- sca(stk, idx, fmodel = fmod2, srmodel = srmod1, qmodel = qmod) 
a4a.stk <- stk + simulate(fit, nsim = 1000)
```

```{r fit-sens1}
## sensitivity 1: updated data truncated to 2018
stk_s <- window(stk, end = 2018)
idx_s <- window(idx, end = 2018)
fit1 <- sca(stk_s, idx_s, fmodel = fmod2020, srmodel = srmod2020, qmodel = qmod) 
a4a.sens1 <- stk_s + simulate(fit1, nsim = 1000)
```

```{r fit-sens2}
## sensitivity 2: updated data truncated to 2018 - old indices
fit2 <- sca(stk_s, idx_2020, fmodel = fmod2020, srmodel = srmod2020, qmodel = qmod2020) 
a4a.sens2 <- stk_s + simulate(fit2, nsim = 1000)
```

```{r comp-plot, fig.height=5, fig.cap=""}
plot(FLStocks(`2026 update` = a4a.stk, `2020 baseline` = a4a.2020))
```

## Model results - 2020 vs sensitivity runs 

```{r sens-plot, fig.height=4, fig.cap="Comparison of the 2020 baseline with the sensitivity runs"}
plot(FLStocks(`2020 baseline` = a4a.2020, `sensitivity 1` = a4a.sens1,
 `sensitivity 2` = a4a.sens2))
```

## Diagnostics (2026 update)
```{r resid, fig.height=4, fig.cap=""}
range(stk)[c('minfbar','maxfbar')] <- c(2,4) 
res <- residuals(fit, stk, idx)
plot(res)
```

## Stock perception: 2020 (left) vs 2026 update (right)
```{r refpts}
### ------------------------------------------------------ ###
# Status - Ref points
### ------------------------------------------------------ ###

stk <- window(swo.stk, start = 1985)
stk <- replaceZeros(stk)
range(stk)[c('minfbar','maxfbar')] <- c(2,4) 
stk <- setPlusGroup(stk, 5) # same plus group
aa.stk <- stk + fit

landings.n(aa.stk)<-catch.n(aa.stk)
landings.wt(aa.stk)<-catch.wt(aa.stk)
landings(aa.stk)<-catch(aa.stk)
discards(aa.stk)<-0
discards.n(aa.stk)<-0
discards.wt(aa.stk)<-0

model = "bevholt"
srr_upd <- fmle(as.FLSR(aa.stk, model = model),control=list(trace=0),fixed=list(b=100))
brp_upd<- brp(FLBRP(aa.stk, sr = srr_upd))
ref_points_upd <- refpts(brp_upd)

aa.stk <- stk_2020 + fit2020
landings.n(aa.stk)<-catch.n(aa.stk)
landings.wt(aa.stk)<-catch.wt(aa.stk)
landings(aa.stk)<-catch(aa.stk)
discards(aa.stk)<-0
discards.n(aa.stk)<-0
discards.wt(aa.stk)<-0
srr_2020 <- fmle(as.FLSR(aa.stk, model = model),control=list(trace=0),fixed=list(b=100))
brp_2020 <- brp(FLBRP(aa.stk, sr = srr_2020))
ref_points_2020 <- refpts(brp_2020)


fmsy_upd <- as.numeric(ref_points_upd['msy','harvest'])
ssbmsy_upd <- as.numeric(ref_points_upd['msy','ssb'])
p1 <- plot(a4a.stk, metrics=list(SSB=ssb, F=fbar)) +
  geom_flpar(data=FLPars(SSB=FLPar(SSBmsy = ssbmsy_upd),
  F=FLPar(FMSY=fmsy_upd)), x=c(1990, 2015))

fmsy_upd <- as.numeric(ref_points_2020['msy','harvest'])
ssbmsy_upd <- as.numeric(ref_points_2020['msy','ssb'])
p2 <- plot(a4a.2020, metrics=list(SSB=ssb, F=fbar)) +
  geom_flpar(data=FLPars(SSB=FLPar(SSBmsy = ssbmsy_upd),
  F=FLPar(FMSY=fmsy_upd)), x=c(1990, 2015))+theme_bw()
```

```{r plotrefpts, fig.cap="",fig.height=4}
library(gridExtra)
grid.arrange(p2, p1, ncol = 2)
```



